Nested sampling with two objective functions

### 2015-11-02 Paper/ms.tex
Random work

\usepackage{dsfont}
\usepackage[utf8]{inputenc}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}

\newcommand{\xx}{\boldsymbol{x}}
\newcommand{\dx}{d\boldsymbol{x}}
\newcommand{\data}{\boldsymbol{D}}
\newcommand{\II}{\boldsymbol{I}}

@@

The corner count of particle $i$ is defined as the number of particles with
a lower value of {\em both} objective functions compared to particle $i$.
Graphically, on a plot of $L_2$ vs. $L_1$ the corner count is the number of
particles in the rectangle whose upper right corner is at
$\left(L_1(\xx_i), L_2(\xx_i)\right)$ (see Figure~\ref{fig:joint2} for an
example). The equation for the corner count is:
\begin{eqnarray}
C_i &=& \sum_{j=1}^N \mathds{1}\left(L_1(\xx_j) < L_1(\xx_i),
L_2(\xx_j) < L_2(\xx_i)
\right).\label{eqn:corner_count}
\end{eqnarray}

@@

We seek an algorithm for computing the partition function
$Z(\beta_1, \beta_2)$ for a range of values of the $\beta$s, from a single
run. To be considered a variant of Nested Sampling, the method should satisfy
the following properties:
\begin{enumerate}
\item The algorithm should begin with $N$ points drawn from the prior $\pi(\xx)$.
\item The algorithm should seek to explore regions where the values of
$L_1(\xx)$ and $L_2(\xx)$ are greater than what would be expected from the
prior.
\item The algorithm should consider a sequence of probability
distributions proportional to $\pi$, but restricted to smaller and smaller
domains for which we can approximately measure the enclosed prior mass.
\item The algorithm should be invariant to monotonic transformations of
$L_1$ and $L_2$, i.e. it should only depend on rankings of $L_1$ and $L_2$
values and not the values themselves.
\end{enumerate}

\begin{figure}
\centering
\includegraphics[scale=0.5]{figures/joint2.pdf}
\caption{\it The same as Figure~\ref{fig:joint1}.
\label{fig:joint2}}
\end{figure}

@@

\begin{algorithm}
\begin{algorithmic}
\State Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$.
\While{more iterations desired}
	\State Calculate the {\it corner count} (Equation~\ref{eqn:corner_count})
of each particle.
	\State Choose one of the particles which has corner count equal to the
minimum nonzero value. If all particles have a corner count of zero,
the algorithm terminates. Call the selected particle particle $K$.
	\State Choose one of the objective functions, and find $K$'s rank in terms
of that objective function.
\EndWhile
\end{algorithmic}
\end{algorithm}

Step 1 of the algorithm ensures that NS property 1 is trivially satisfied.

@@

\begin{figure}
\centering
\includegraphics[scale=0.75]{figures/truth.pdf}
\caption{\label{fig:truth}}
\end{figure}

>>>>

\usepackage{dsfont}
\usepackage[utf8]{inputenc}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{algorithmicx}
\usepackage{algpseudocode}

\DeclareUnicodeCharacter{00A0}{ }
\newcommand{\xx}{\boldsymbol{x}}
\newcommand{\dx}{d\boldsymbol{x}}
\newcommand{\data}{\boldsymbol{D}}
\newcommand{\II}{\boldsymbol{I}}

@@

The corner count of particle $i$ is defined as the number of particles with
a lower value of {\em both} objective functions compared to particle $i$.
Graphically, on a plot of $L_2$ vs. $L_1$ the corner count is the number of
particles in the rectangle whose upper right corner is at
$\left(L_1(\xx_i), L_2(\xx_i)\right)$ (see Figure~\ref{fig:joint2} for an
example). The equation for the corner count of particle $i$ is:
\begin{eqnarray}
C_i &=& \sum_{\substack{j=1\\{j \neq i}}}^N \mathds{1}\left(L_1(\xx_j) < L_1(\xx_i),
L_2(\xx_j) < L_2(\xx_i)
\right).\label{eqn:corner_count}
\end{eqnarray}

@@

We seek an algorithm for computing the partition function
$Z(\beta_1, \beta_2)$ for a range of values of the $\beta$s, from a single
run. To be considered a variant of Nested Sampling, the method should satisfy
the following properties:
\begin{enumerate}
\item The algorithm should begin with $N$ points drawn from the prior $\pi(\xx)$.
\item The algorithm should seek to explore regions where the values of
$L_1(\xx)$ and $L_2(\xx)$ are higher than the prior would typically predict.
\item The algorithm should consider a sequence of probability
distributions proportional to $\pi$, but restricted to smaller and smaller
domains for which we can approximately measure the enclosed prior mass.
\item The algorithm should be invariant to monotonic transformations of
$L_1$ and $L_2$, i.e. it should only depend on rankings of $L_1$ and $L_2$
values and not the values themselves.
\end{enumerate}
The standard Nested Sampling algorithm, (as well as alternatives such as
Diffusive Nested Sampling \citep{dnest}) has these properties.

@@

\begin{algorithm}
\begin{algorithmic}
\State Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$, and
initialise the {\em remaining prior mass} at $M=1$.
\While{more iterations desired}
	\State Calculate the {\it corner count} (Equation~\ref{eqn:corner_count})
of each particle.
	\State Choose one of the particles which has corner count equal to the
minimum nonzero value. Label this particle number $K$.
	\State Choose one of the objective functions at random. Label this
objective function $J$. Let $L_{J}(\xx_K)$ be the value of this objective
function for particle $K$.
\EndWhile
\end{algorithmic}
\end{algorithm}

Step 1 of the algorithm ensures that NS property 1 is trivially satisfied.
In step 2, we try to find a particle with the lowest possible nonzero
corner count. Often this will be 1. In Figure~\ref{fig:joint2}, particle $K$
is located at $(2.65, 1.31)$ and has a corner count of 1.

At step 2, if all particles have a corner count of 0, the algorithm terminates.

\begin{figure}
\centering
\includegraphics[scale=0.5]{figures/joint2.pdf}
\caption{\it The same as Figure~\ref{fig:joint1}. The selected particle
at $(2.65, 1.31)$ has a corner count of 1. The particle at
$(2.50, 0.84)$ can be thought of as a single-particle standard Nested Sampling
run restricted to $L_1 < 2.65$ which is attempting to increase $L_1$.
\label{fig:joint2}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=0.75]{figures/truth.pdf}
\caption{The true partition function and KL divergence (from the prior to
the canonical distribution) as a function of two temperatures for the
test problem.\label{fig:truth}}
\end{figure}

\begin{figure}
\centering
\includegraphics[scale=0.75]{figures/output.png}
\caption{\label{fig:output}}
\end{figure}
